---
title: "Time-RSA Analysis"
output:
  html_document:
    df_print: paged
---

```{r}
library(tidyverse)
library(lme4)
library(lmerTest)
library(emmeans)
```

```{r config}
# Path to the long-format comparison table produced by:
#     python run_pipeline.py compare
# Override it here if your output lives elsewhere (e.g. RSA_PLANNING_OUTPUT).
comparison_csv <- "../output/comparisons/comparison_long.csv"

stopifnot(file.exists(comparison_csv))
```

```{r}
data <- read.csv(comparison_csv) %>%
  separate(tc_id, into = c("lab", "id", "numcrit", "awar")) %>%
  unite(tc_id, id, numcrit) %>%
  select(-lab, -awar)
```

# Analysis for WHEN:

## 1. Proportion of Missed Critical: \# missed_critical / totalnum_critical

a. COMBO of both planning1:belief_tracking1 -- strong reduction, better than each individually
b. both significant, but: belief_tracking1 (myopic) > planning (uniform)


```{r}
for_when1 <- data %>%
  select(tc_id, dispersion = density, num_critical_prop, sharing_metric, user_general_aware, user_critical_aware, model_name, prop_missed_critical) %>%
  mutate(user_general_aware = 
           factor(user_general_aware, levels = c("unaware", "aware")),
         user_critical_aware = 
           user_critical_aware %>% str_remove("%") %>% as.numeric(),
         planning = if_else(model_name %in% c("full", "uniform"), "yes", "no") %>% factor(levels = c("no", "yes")),
         belief_tracking = if_else(model_name %in% c("full", "myopic"), "yes", "no") %>% factor(levels = c("no", "yes")),
         dispersion.c = scale(dispersion, 
                           center=TRUE, scale=FALSE) %>% as.vector(),
         num_critical_prop.c = scale(num_critical_prop, 
                                   center=TRUE, scale=FALSE)  %>% as.vector(),
         sharing_metric.c = scale(sharing_metric, 
                                center=TRUE, scale=FALSE) %>% as.vector(),
         user_critical_aware.c = scale(user_critical_aware,
                                       center=TRUE, scale=FALSE) %>% as.vector())

contrasts(for_when1$user_general_aware) <- c(-1, 1)
contrasts(for_when1$planning) <- c(-1, 1)
contrasts(for_when1$belief_tracking) <- c(-1, 1)
```

```{r}
lm_when1 <- lmer(prop_missed_critical ~ planning + belief_tracking + planning:belief_tracking +
                num_critical_prop.c + dispersion.c + sharing_metric.c + 
                user_critical_aware.c + user_general_aware +
                planning:num_critical_prop.c + 
                planning:dispersion.c + 
                planning:sharing_metric.c +
                planning:user_critical_aware.c + 
                planning:user_general_aware +
                belief_tracking:num_critical_prop.c + 
                belief_tracking:dispersion.c + 
                belief_tracking:sharing_metric.c +
                belief_tracking:user_critical_aware.c + 
                belief_tracking:user_general_aware +
                planning:belief_tracking:num_critical_prop.c + 
                planning:belief_tracking:dispersion.c + 
                planning:belief_tracking:sharing_metric.c +
                planning:belief_tracking:user_critical_aware.c + 
                planning:belief_tracking:user_general_aware +
                num_critical_prop.c:dispersion.c +
                num_critical_prop.c:sharing_metric.c +
                num_critical_prop.c:user_critical_aware.c +
                num_critical_prop.c:user_general_aware +
                dispersion.c:sharing_metric.c +
                dispersion.c:user_critical_aware.c +
                dispersion.c:user_general_aware +
                sharing_metric.c:user_critical_aware.c +
                sharing_metric.c:user_general_aware +
                user_critical_aware.c:user_general_aware +
              (1 | tc_id),
              data=for_when1)

summary(lm_when1)
```

```{r}
pairwise_models <- emmeans(lm_when1, pairwise ~ planning*belief_tracking, pbkrtest.limit = 3200)
summary(pairwise_models)
```

## 2. Proportion of Missed Critical Properties which were Low Awareness: \# missed_low_critical / \# critical low aware

Similar trends as above, but the benefit of planning seems to be further low: no sign. difference between Baseline and Uniform (although this is better ofc)!

```{r}
for_when2 <- data %>%
  select(tc_id, dispersion = density, num_critical_prop, sharing_metric, user_general_aware, user_critical_aware, model_name, prop_missed_low_critical) %>%
  mutate(user_general_aware = 
           factor(user_general_aware, levels = c("unaware", "aware")),
         user_critical_aware = 
           user_critical_aware %>% str_remove("%") %>% as.numeric(),
         planning = if_else(model_name %in% c("full", "uniform"), "yes", "no") %>% factor(levels = c("no", "yes")),
         belief_tracking = if_else(model_name %in% c("full", "myopic"), "yes", "no") %>% factor(levels = c("no", "yes")),
         dispersion.c = scale(dispersion, 
                           center=TRUE, scale=FALSE) %>% as.vector(),
         num_critical_prop.c = scale(num_critical_prop, 
                                   center=TRUE, scale=FALSE)  %>% as.vector(),
         sharing_metric.c = scale(sharing_metric, 
                                center=TRUE, scale=FALSE) %>% as.vector(),
         user_critical_aware.c = scale(user_critical_aware,
                                       center=TRUE, scale=FALSE) %>% as.vector())

contrasts(for_when2$user_general_aware) <- c(-1, 1)
contrasts(for_when2$planning) <- c(-1, 1)
contrasts(for_when2$belief_tracking) <- c(-1, 1)

lm_when2 <- lmer(prop_missed_low_critical ~ planning + belief_tracking + planning:belief_tracking +
                num_critical_prop.c + dispersion.c + sharing_metric.c + 
                user_critical_aware.c + user_general_aware +
                planning:num_critical_prop.c + 
                planning:dispersion.c + 
                planning:sharing_metric.c +
                planning:user_critical_aware.c + 
                planning:user_general_aware +
                belief_tracking:num_critical_prop.c + 
                belief_tracking:dispersion.c + 
                belief_tracking:sharing_metric.c +
                belief_tracking:user_critical_aware.c + 
                belief_tracking:user_general_aware +
                planning:belief_tracking:num_critical_prop.c + 
                planning:belief_tracking:dispersion.c + 
                planning:belief_tracking:sharing_metric.c +
                planning:belief_tracking:user_critical_aware.c + 
                planning:belief_tracking:user_general_aware +
                num_critical_prop.c:dispersion.c +
                num_critical_prop.c:sharing_metric.c +
                num_critical_prop.c:user_critical_aware.c +
                num_critical_prop.c:user_general_aware +
                dispersion.c:sharing_metric.c +
                dispersion.c:user_critical_aware.c +
                dispersion.c:user_general_aware +
                sharing_metric.c:user_critical_aware.c +
                sharing_metric.c:user_general_aware +
                user_critical_aware.c:user_general_aware +
              (1 | tc_id),
              data=for_when2)

summary(lm_when2)
```

```{r}
pairwise_models <- emmeans(lm_when2, pairwise ~ planning*belief_tracking, pbkrtest.limit = 3200)
summary(pairwise_models)
```

## 3. Time to First Alert about a Low Awareness Critical Property:

both decrease, but that due to belief tracking is larger.

```{r}
for_when3 <- data %>%
  select(tc_id, dispersion = density, num_critical_prop, sharing_metric, user_general_aware, user_critical_aware, model_name, med_delay_low) %>%
  # Just to be safe removing instances where low awareness situations didn't occur
  filter(med_delay_low != -1) %>%
  mutate(user_general_aware = 
           factor(user_general_aware, levels = c("unaware", "aware")),
         user_critical_aware = 
           user_critical_aware %>% str_remove("%") %>% as.numeric(),
         planning = if_else(model_name %in% c("full", "uniform"), "yes", "no") %>% factor(levels = c("no", "yes")),
         belief_tracking = if_else(model_name %in% c("full", "myopic"), "yes", "no") %>% factor(levels = c("no", "yes")),
         dispersion.c = scale(dispersion, 
                           center=TRUE, scale=FALSE) %>% as.vector(),
         num_critical_prop.c = scale(num_critical_prop, 
                                   center=TRUE, scale=FALSE)  %>% as.vector(),
         sharing_metric.c = scale(sharing_metric, 
                                center=TRUE, scale=FALSE) %>% as.vector(),
         user_critical_aware.c = scale(user_critical_aware,
                                       center=TRUE, scale=FALSE) %>% as.vector())


contrasts(for_when3$user_general_aware) <- c(-1, 1)
contrasts(for_when3$planning) <- c(-1, 1)
contrasts(for_when3$belief_tracking) <- c(-1, 1)

lm_when3 <- lmer(med_delay_low ~ planning + belief_tracking + planning:belief_tracking +
                num_critical_prop.c + dispersion.c + sharing_metric.c + 
                user_critical_aware.c + user_general_aware +
                planning:num_critical_prop.c + 
                planning:dispersion.c + 
                planning:sharing_metric.c +
                planning:user_critical_aware.c + 
                planning:user_general_aware +
                belief_tracking:num_critical_prop.c + 
                belief_tracking:dispersion.c + 
                belief_tracking:sharing_metric.c +
                belief_tracking:user_critical_aware.c + 
                belief_tracking:user_general_aware +
                planning:belief_tracking:num_critical_prop.c + 
                planning:belief_tracking:dispersion.c + 
                planning:belief_tracking:sharing_metric.c +
                planning:belief_tracking:user_critical_aware.c + 
                planning:belief_tracking:user_general_aware +
                num_critical_prop.c:dispersion.c +
                num_critical_prop.c:sharing_metric.c +
                num_critical_prop.c:user_critical_aware.c +
                num_critical_prop.c:user_general_aware +
                dispersion.c:sharing_metric.c +
                dispersion.c:user_critical_aware.c +
                dispersion.c:user_general_aware +
                sharing_metric.c:user_critical_aware.c +
                sharing_metric.c:user_general_aware +
                user_critical_aware.c:user_general_aware +
              (1 | tc_id),
              data=for_when3)

summary(lm_when3)
```

```{r}
pairwise_models <- emmeans(lm_when3, pairwise ~ planning*belief_tracking)
summary(pairwise_models)
```

## 4. Prioritization Index

### model prioritizes alerting about low-awareness critical properties faster - less delay? (+ve: Low-awareness; -ve high-aware/other; 0: discard)

BOTH uniform and baseline unable to prioritize low awareness properties, as would be expected.
Myopic seems to not prioritize, however doesn't have good confidence.
Only the full model can actually show statistically significant preference for low awareness crit. props


```{r}
for_when4 <- data %>%
  select(tc_id, dispersion = density, num_critical_prop, sharing_metric, user_general_aware, user_critical_aware, model_name, prioritization_index) %>%
  mutate(user_general_aware = 
           factor(user_general_aware, levels = c("unaware", "aware")),
         user_critical_aware = 
           user_critical_aware %>% str_remove("%") %>% as.numeric(),
         planning = if_else(model_name %in% c("full", "uniform"), "yes", "no") %>% factor(levels = c("no", "yes")),
         belief_tracking = if_else(model_name %in% c("full", "myopic"), "yes", "no") %>% factor(levels = c("no", "yes")),
         dispersion.c = scale(dispersion, 
                           center=TRUE, scale=FALSE) %>% as.vector(),
         num_critical_prop.c = scale(num_critical_prop, 
                                   center=TRUE, scale=FALSE)  %>% as.vector(),
         sharing_metric.c = scale(sharing_metric, 
                                center=TRUE, scale=FALSE) %>% as.vector(),
         user_critical_aware.c = scale(user_critical_aware,
                                       center=TRUE, scale=FALSE) %>% as.vector())

contrasts(for_when4$user_general_aware) <- c(-1, 1)
contrasts(for_when4$planning) <- c(-1, 1)
contrasts(for_when4$belief_tracking) <- c(-1, 1)

lm_when4 <- lmer(prioritization_index ~ planning + belief_tracking + planning:belief_tracking +
                num_critical_prop.c + dispersion.c + sharing_metric.c + 
                user_critical_aware.c + user_general_aware +
                planning:num_critical_prop.c + 
                planning:dispersion.c + 
                planning:sharing_metric.c +
                planning:user_critical_aware.c + 
                planning:user_general_aware +
                belief_tracking:num_critical_prop.c + 
                belief_tracking:dispersion.c + 
                belief_tracking:sharing_metric.c +
                belief_tracking:user_critical_aware.c + 
                belief_tracking:user_general_aware +
                planning:belief_tracking:num_critical_prop.c + 
                planning:belief_tracking:dispersion.c + 
                planning:belief_tracking:sharing_metric.c +
                planning:belief_tracking:user_critical_aware.c + 
                planning:belief_tracking:user_general_aware +
                num_critical_prop.c:dispersion.c +
                num_critical_prop.c:sharing_metric.c +
                num_critical_prop.c:user_critical_aware.c +
                num_critical_prop.c:user_general_aware +
                dispersion.c:sharing_metric.c +
                dispersion.c:user_critical_aware.c +
                dispersion.c:user_general_aware +
                sharing_metric.c:user_critical_aware.c +
                sharing_metric.c:user_general_aware +
                user_critical_aware.c:user_general_aware +
              (1 | tc_id),
              data=for_when4)

summary(lm_when4)
```

```{r}
pairwise_models <- emmeans(lm_when4, pairwise ~ planning*belief_tracking, pbkrtest.limit = 3200)
summary(pairwise_models)
```
